---
title: "Phylogeny9"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: darkly
    highlight: rstudio
    code_folding: show
    df_print: paged
    toc: true
    toc_depth: 2
    toc_float: true
---

# Parameters
```{r}
path<-"Z:/EQUIPE/meiorec/PROJECTS/TransciptomeScreen/Phylo/Data/"
template_tree_path<-"Z:/EQUIPE/meiorec/PROJECTS/TransciptomeScreen/Phylo/Data/"
template_tree_file<-"Viridiplantae2.txt"
```

```{r}
E_value_threshold<-as.numeric(10e-10)
Identity_threshold<-10
Iteration<-"2"
n_sequences<-10000
tree_title<-"Tree submitted for paper"
files_vector<-c("SCEP1_PSI_BLAST_10000_viridi_iteration2_Descriptions.csv",
                "SCEP2_PSI_BLAST_10000_viridi_iteration2_Descriptions.csv",
                "ZYP1a_PSI_BLAST_10000_viridi_iteration1_Descriptions.csv")

ordered_names<-c("SCEP1",
                 "SCEP2",
                 "ZYP1")
```


# Load libraries
```{r, message=FALSE, warning=FALSE}
library(tidyr)
library(tidyverse)
library(dplyr)
library(taxize)
library(rentrez)
library("treeio")
library(ggplot2)
library(ggtree)
library(dplyr)
library(tidytree)
library(ape)
library(data.table)
library(useful)
library(stringr)
library(ggrepel)
#library(peuchmar)

outersect <- function(x, y) {
  sort(c(setdiff(x, y),
         setdiff(y, x)))
}

get_lines <-function(df, vector, df_column){
  len<-nrow(df)
  result_df <- data.frame(matrix(ncol = ncol(df), nrow = 0))
  colnames(result_df) <- colnames(df)
  for (i in 1:len){
    if (any(vector  %in% df_column[i])==TRUE){
      result_df<-rbind(result_df,df[i,])
    }
  }
  print(result_df)
}

```
# Load Data

NCBI downloaded data description table

```{r}
df<-data.frame()
for (i in c(1:length(files_vector))){
  file<-read.csv(paste(path, files_vector[i], sep = ""), header=TRUE)
  file$file<-ordered_names[i]
  df<-rbind(df, file)
}
```

# Filtering data
```{r}
df$E.value<-as.numeric(df$E.value)
df$Per..ident<-as.numeric(df$Per..ident)
df_raw<-df

df<-df %>% filter(E.value < E_value_threshold) %>% filter(Per..ident > Identity_threshold)
df[order(df$file,df$Scientific.Name),]

tail(df)

```

```{r}
species_to_change<-c("Vigna angularis var. angularis",
                     "Perilla frutescens var. hirtella",
                     "Perilla frutescens var. frutescens",
                     "Panicum hallii var. hallii",
                     "Oryza sativa Indica Group",
                     "Olea europaea var. sylvestris",
                     "Olea europaea subsp. europaea",
                     "Hordeum vulgare subsp. vulgare",
                     "Cucurbita argyrosperma subsp. sororia",
                     "Cucurbita argyrosperma subsp. argyrosperma",
                     "Cucumis melo var. makuwa",
                     "Camellia sinensis var. sinensis",
                     "Brassica rapa subsp. trilocularis",
                     "Brassica oleracea var. oleracea",
                     "Arabidopsis lyrata subsp. lyrata",
                     "Ananas comosus var. bracteatus",
                     "Oryza sativa Japonica Group",
                     "Beta vulgaris subsp. vulgaris",
                     "Daucus carota subsp. sativus",
                     "Musa acuminata subsp. malaccensis",
                     "Physcomitrium patens",
                     "Volvox carteri f. nagariensis",
                     "Coccomyxa subellipsoidea C-169",
                     "Micromonas pusilla CCMP1545 ")

species_to_change_to<-c("Vigna angularis",
                     "Perilla frutescens",
                     "Perilla frutescens",
                     "Panicum hallii",
                     "Oryza sativa",
                     "Olea europaea",
                     "Olea europaea",
                     "Hordeum vulgare",
                     "Cucurbita argyrosperma",
                     "Cucurbita argyrosperma",
                     "Cucumis melo",
                     "Camellia sinensis",
                     "Brassica rapa",
                     "Brassica oleracea",
                     "Arabidopsis lyrata",
                     "Ananas comosus",
                     "Oryza sativa",
                     "Beta vulgaris",
                     "Daucus carota",
                     "Musa acuminata",
                     "Physcomitrella patens",
                     "Volvox carteri",
                     "Coccomyxa subellipsoidea",
                     "Micromonas pusilla")
```

```{r}
path_phylo3<-template_tree_path
plants<-read.newick(paste(path_phylo3, tree_file, sep = ""))
plants$tip.label<- sub("_", " ", plants$tip.label)

for (i in 1:length(species_to_change)){
  if (length(species_to_change)!=length(species_to_change_to)){
    warning("The two character vectors have different lengths !")
  }
  df$Scientific.Name[df$Scientific.Name == species_to_change[i]] <- species_to_change_to[i]
} 

df_homologs<-get_lines(df, plants$tip.label, df$Scientific.Name)

write.csv(df_homologs,paste(path,"species_homologs.txt", sep = ""), row.names = FALSE)
```


# Exploring data

```{r}
for (i in c(1:length(files_vector))){
species_summary<-get_lines(df,c(ordered_names[i]),df$file) %>% 
  group_by(Scientific.Name) %>%
  summarise(no_rows = length(Scientific.Name))
names(species_summary)<- c('Scientific.Name', paste0("number_",ordered_names[i]))
print(species_summary)
}
```

Look into specific species:
```{r}
sp<-c("Selaginella moellendorffii")
specific_sp<-get_lines(df,sp,df$Scientific.Name)
```

# Load tree

Loading tree (here tree taken from Fernandes et al. 2018, subset viridiplantae):
```{r}
plants<-read.newick(paste(path,template_tree_file, sep = ""))
```

# Exploring tree 

## Full tree

Visualizing tree data:
```{r}
plants$tip.label<- sub("_", " ", plants$tip.label)
print(n=250,as_tibble(plants))
```


# Comparing species set

Vector containing all the species in the plant tree:
```{r}
plant_species<-plants$tip.label
plant_species
```


```{r}
scep1<-get_lines(df, c("SCEP1"), df$file)
scep1_sp_list<-scep1$Scientific.Name

scep2<-get_lines(df, c("SCEP2"), df$file)
scep2_sp_list<-scep2$Scientific.Name

zyp1<-get_lines(df, c("ZYP1"), df$file)
zyp1_sp_list<-zyp1$Scientific.Name
```

Intersection between all proteins and plant tree:
```{r}
intersecting_species_1<-intersect(plant_species,scep1_sp_list)
intersecting_species_2<-intersect(plant_species,scep2_sp_list)
intersecting_species_3<-intersect(plant_species,zyp1_sp_list)

intersecting_species_12<-intersect(intersecting_species_1,intersecting_species_2)
intersecting_species_123<-intersect(intersecting_species_12,intersecting_species_3)
intersecting_species_full<-intersect(intersecting_species_123,intersecting_species_z)

```

# Annotating tree 

Producing annotation matrix:
```{r}
df_for_output<-df
number_of_homologs<-table(df$Scientific.Name, df$file)
df_tree_species<-get_lines(df, plant_species, df$Scientific.Name)
number_of_homologs_from_tree<-table(df_tree_species$Scientific.Name, df_tree_species$file)
sp_mat<-as.matrix(number_of_homologs_from_tree)
sp_mat[sp_mat>0] <- "present"
sp_mat[sp_mat==0] <- "absent"
sp_mat
```
Adding fully absent lines
```{r}
row.names(sp_mat)
outs<-outersect(plant_species,row.names(sp_mat))
miss_mat<-matrix(,nrow=length(outs), ncol = length(ordered_names))
row.names(miss_mat)<-outs
miss_mat[]<-"absent"
full_mat<-rbind(sp_mat,miss_mat)
df<-data.frame(full_mat)
df
```


```{r}
p1_1<-ggtree(plants, color="black",
           layout = "rectangular",
           ladderize = TRUE)+
   #geom_tiplab(geom='label',
    #          color="black",
     #         size=2.2)+
  #geom_nodepoint(color="black", alpha=1, size=0)+
  ggplot2::xlim(0, 40)
  
  #ggplot2::ylim(0, 50)+
  #ylim(0,50)
  #geom_tippoint(color="#FDAC4F", shape=8, size=2)

```
# Drawing tree and heatmap

```{r}
p1<-p1_1 %<+% df + geom_tiplab(
                        geom = "text",
                        size=7)+
  ggtitle(paste(tree_title,
                paste(paste("E.value<",E_value_threshold, sep=""),
                      paste(paste("%identity>",Identity_threshold, sep=""),
                            paste(paste("n=",n_sequences, sep=""), 
                                  paste("iteration", Iteration),sep=", "),sep=", "),sep=", "),sep=", "))
```


```{r}

gheatmap(p1, df, offset=13, width=0.6, font.size=7, 
         colnames_angle=0, hjust=0, 
         colnames=TRUE, color="azure", colnames_position="top",
         colnames_offset_x = -1) +
  scale_fill_manual(breaks=c("present", "absent", "Unsure"), 
                    values=c("dodgerblue1", "firebrick1", "bisque"), name="Homologs")+
  theme(legend.key.size = unit(1, 'cm'), #change legend key size
        legend.key.height = unit(1, 'cm'), #change legend key height*
        legend.key.width = unit(1, 'cm'), #change legend key width
        legend.title = element_text(size=20), #change legend title font size
        legend.text = element_text(size=18))+ #change legend text font size
  scale_x_ggtree()

```

```{r,eval=FALSE}
ggsave(filename = "ncbi_blast_intersect_tree.svg" , device = "svg", 
       width = 20, height = 18 , units = "in" ,
       limitsize = FALSE, path = path)
```

# SupTable

```{r}
df_for_output<-get_lines(df_for_output,plant_species, df_for_output$Scientific.Name)

write.csv2(df_for_output, paste(path,"Homologs_SupTable.csv", sep=""), row.names = FALSE)
```

